---
title: "Required packages"
author: "Felipe Walter Pereira"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Load the required packages

```{r}
packages <- c("devtools", "raster", "rgdal", "rgeos", "scales", "spThin")
req_pack <- packages[!(packages %in% installed.packages()[, "Package"])]
```

Install the required packages if necessary

```{r}
if (length(req_pack) > 0) {install.packages(req_pack, dependencies = TRUE)}
```

The following chunk fix 'rgdal' and 'rgeos' dependencies

```{r}
rgeos_path <- "https://cran.r-project.org/src/contrib/Archive/rgeos/rgeos_0.6-4.tar.gz"
if(!require("rgeos")){
  install.packages(rgeos_path,repos = NULL,type = "source")
}
rgdal_path <- "https://cran.r-project.org/src/contrib/Archive/rgdal/rgdal_1.6-7.tar.gz"
if(!require("rgdal")){
  install.packages(rgdal_path, repos = NULL,type = "source")
}
```


'grinnell' and 'kuenm' R packages installation (both require 'rgeos' and 'rgdal')

```{r}
devtools::install_github("fmachados/grinnell")
library("grinnell")
devtools::install_github("marlonecobos/kuenm")
library("kuenm")
```

## renv - managing the packages versions

```{r}
install.packages("renv")
library(renv)
```

Copying packages into project library

```{r}
renv::init()
```

Checking the dependences status

```{r}
renv::status()
```

Use the following function whenever install or update a package - it ensures that the package version will be registered in the renv library

```{r}
renv::snapshot()
```